The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2012-10-24 bmc_article.tex
Added Ola as co-author
Writing some content

\author{
  Egon L Willighagen$^{1}$% // FIXED FIRST AUTHOR
  \email{Egon L Willighagen - egon.willighagen@ki.se}
\and
  Miguel Rojas Cherto$^2$%
  \email{Miguel Rojas Cherto - steinbeck@ebi.ac.uk}
\and
  You C B Here$^3$%
  \email{You C B Here - you.here@example.org}
\and
  C D K Project\correspondingauthor$^98$% // FIXED SECOND BUT LAST AUTHOR
  \email{C D K Project - cdk-devel@lists.sourceforge.net}
\and
  Rajarshi Guha$^99$% // FIXED ONE BUT LAST AUTHOR
  \email{Rajarshi Guha - steinbeck@ebi.ac.uk}
\and
  Christoph Steinbeck$^100$% // FIXED LAST AUTHOR
  \email{Christoph Steinbeck - steinbeck@ebi.ac.uk}
}

@@

Recent prominent use cases of the CDK~\cite{Steinbeck2003,Steinbeck2006} \ldots (EGON)

@@

  The CDK has adopted a new atom typing algorithm \ldots (EGON)

@@

  \subsubsection*{InChI integration}

  .... jni-inchi dep, atom numbering, tautomer generation

@@

%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}
  Text for this section \ldots

>>>>

\author{
  Egon L Willighagen$^{1}$% // FIXED FIRST AUTHOR
  \email{Egon L Willighagen - egon.willighagen@ki.se}
\and
  Miguel Rojas Cherto$^2$%
  \email{Miguel Rojas Cherto - steinbeck@ebi.ac.uk}
\and
  Ola Spjuth$^3$%
  \email{Ola Spjuth - ola.spjuth@farmbio.uu.se}
\and
  You C B Here$^4$%
  \email{You C B Here - you.here@example.org}
\and
  C D K Project\correspondingauthor$^98$% // FIXED SECOND BUT LAST AUTHOR
  \email{C D K Project - cdk-devel@lists.sourceforge.net}
\and
  Rajarshi Guha$^{99}$% // FIXED ONE BUT LAST AUTHOR
  \email{Rajarshi Guha - steinbeck@ebi.ac.uk}
\and
  Christoph Steinbeck$^{100}$% // FIXED LAST AUTHOR
  \email{Christoph Steinbeck - steinbeck@ebi.ac.uk}
}

@@

Recent prominent use cases of the CDK~\cite{Steinbeck2003,Steinbeck2006} \ldots (EGON)
AMBIT, ScaffoldHunter, OMG, ...

@@

  The CDK has adopted a new atom typing framework, to split out the perception of atom types
  from the algorithms that use atom type properties. Previous CDK releases had perception
  integrated with algorithms, even though there was only one set of defined atom types.
  This led to alternative but contradiction perception algorithms, which made maintenance
  of these code bases impractical. For example, adding a new atom type required all alternatives
  to be updated.
  To solve this, a single perception class was added, called CDKAtomTypeMatcher. The actual
  algorithm used is based
  on heuristic rules, and alternative approach for perception exist. Possibly, in the future
  alternative algorithms may be implemented, while that would again increase the maintenance
  burden.

  The new class can perceive the atom type of single atoms, but also for a full atom container.
  The CDKAtomTypeMatcher class is written to handle various amounts of missing information,
  and take into account as much information from the input. Perception

  \subsubsection*{Stereochemistry}

  CDK 1.4.XXXX [CHECK] introduces a new API for stereochemistry information: the interfaces
  IStereoChemistryElement was introduced with

@@

@@

\subsection*{Binary distributions}

\subsubsection*{Maven packages}

INVITE NINA

\subsubsection*{OSGi bundles}

INVITE ARVID

%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}

Over the past years since the last CDK publication in 2006, a lot has changed. The functionality
has seen many additions, and the stability of the development model has significantly improved.
This paper outlines how peer review and quality control contributed to a much wider adoption
of our cheminformatics library and seen many new additions. With over 75 contributors,
the CDK is alive and kicking.

However, there are many unsolved issues. Performance is perhaps one of the most important ones.
The disadvantage of having an API that supports many data models, is that the API is heavy and
the interface implementations hard to optimize.
